---
title: "figure_2_sal_paper"
author: "Joao Carlos Gomes-Neto"
date: "1/6/2021"
output: html_document
---

Loading the necessary packages 
```{r, include = FALSE}
# Load packages

library(tidyverse)
library(skimr)
library(RColorBrewer)
library(gridExtra)
library(lattice)
library(ggpubr)
library(vegan)
library(reshape2)
library(ggrepel)
library(ggnewscale)
library(forcats)
library(naniar)
library(Rtsne)

# Install ggtree 
#if (!requireNamespace("BiocManager", quietly = TRUE))
  # install.packages("BiocManager")
#BiocManager::install(version = "3.11")

#BiocManager::install("ggtree")

# Load ggtree

library(ggtree)
```

Salmonella Infantis phylogeny

```{r, include = FALSE}
# Enter the phylogeny for Typhimurium group1

tree <- read.tree("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/infantis/input_data/infantis_phylogeny.tree")

#############################################

# Enter Baps 1-6 

baps <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/infantis/input_data/fastbaps_partition_baps_prior_l6.csv')

# Changing column names 

colnames(baps)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/infantis/input_data/salmonellast_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'_'), "[", 1)

# Select columns of interest

s1 <- mlst %>%
          select(id2, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") %>%
                  rename(id = id2)

# QC the data

skim(s1)

####################################################################

# Enter SISTR results

sistr <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/infantis/input_data/sistr_all.csv')

# QC the data 

skim(sistr)

# Generate the id column 

sistr$id <- sapply(strsplit(as.character(sistr$genome),'_'), "[", 1)

# Select columns 

sis1 <- sistr %>%
            select(id, cgmlst_ST, serovar_cgmlst) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") 

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mslt first

d1 <- left_join(b1, s1, on = "id")
d2 <- left_join(d1, sis1, on = "id")

# Create new columns for ST, cgmlst, and serovar

# cgmlst

a <- d2 %>%
  group_by(ST, cgmlst_ST) %>%
  count() %>%
  arrange(desc(n))


d2 <- mutate(d2, cgmlst = ifelse(cgmlst_ST %in% 2242423463, "cgMLST 2242423463",
                                            "Other cgMLSTs"))

# st

d2$st <- ifelse(d2$ST == 32, "ST32", "Other STs")

# Replace NA in the ST column

d2 <- d2 %>% mutate(st = replace_na(st, "Other STs"))

# Create serovar column

#data1 <- data1 %>% mutate(serovar_cgmlst = replace_na(serovar_cgmlst, 0))
d2$serovar <- ifelse(d2$serovar_cgmlst == "Infantis", "Infantis",
                            "Other serovars")
                     
# Check to make sure the relevant columns are ok

table(d2$st)
table(d2$cgmlst)
table(d2$serovar)
table(d2$baps_1)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d3 <- d2 %>% 
                    select("id", "serovar", "baps_1", "st", "cgmlst") %>%
                    mutate(baps_1 = as.factor(baps_1))

# Make id the rowname or index

d3 <- column_to_rownames(d3, var = "id") 

# Create the tree

i1 <- ggtree(tree, layout = "circular") + xlim(-50, NA)


fig4a <- gheatmap(i1, d3, offset=.0, width=10, colnames = FALSE) +
  scale_fill_manual(values = c("darkblue", "gray",
                               "brown", "purple", 
                                 "darkgreen", "red",
                               "gray", "black", "orange"),
                    breaks = c("Infantis", "Other serovars",
                               "1", "2", "3",
                               "ST32", "Other STs",
                               "cgMLST 2242423463", "Other cgMLSTs"),
                    name="Serovar (innermost) -> BAPS1 -> ST -> cgMLST (outermost)") +
    ggtitle(expression(bold(paste("A. Hierarchical population structure - ", bolditalic(S),". Infantis")))) +
  theme(plot.title = element_text(size = 40, face = "bold"), 
        legend.title=element_text(size=26, face = "bold"), 
    legend.text=element_text(size=22))
fig4a
```

Salmonella infantis - distribution of SNPs 

```{r, include = FALSE}

# enter the snp data

data1 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/infantis/input_data/distance_snp_sites_infantis_columns.csv')

# filter out distance between itself

data2 <- data1 %>% filter_(~id_1 != id_2)

# distribution of snps 

summary(data2$dist)

# Generate quantiles

data3 <- data2 %>% 
             summarize(quartile = quantile(dist, probs = c(0.25, 0.5, 0.75, 1)))
data3

# creating a quartile variable for distribution of snps

data2 <- data2 %>% mutate(quartile = case_when(dist <= 36 ~ "quartile 1",
                                                 dist <= 122 ~ "quartile 2",
                                                 dist <= 157 ~ "quartile 3",
                                                 dist > 127368 ~ "quartile 4"))

# Plot the distribution of SNPs

snpinf <- ggplot(data2, aes(x = dist)) +
  geom_histogram(bins = 10, fill = "darkblue") +
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("Number of SNPs") +
  ylab("Frequency") +
  theme(axis.text.y = element_text(size = 40)) +
  theme(axis.title.y = element_text(size = 45, face = "bold")) +
  theme(axis.title.x = element_text(size = 45, face = "bold")) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.7, size = 40)) +
  ggtitle("") +
  theme(plot.title = element_text(size = 50, face = "bold")) + 
  annotate(geom = "text", label = "Mean = 723", x = 5e+04, y = 6e+06, 
                                  color = "black", size = 18) +
  annotate(geom = "text", label = "Median = 122", x = 5e+04, y = 5e+06, 
                                  color = "black", size = 18) +
  annotate(geom = "text", label = "SD = 5,385", x = 5e+04, y = 4e+06, 
                                  color = "black", size = 18) +
  annotate(geom = "text", label = "Range (0-127,368)", x = 5e+04, y = 3e+06, 
                                  color = "black", size = 18) 
snpinf

# get stats on snp dist

snpstats <- data2 %>% 
           summarize(mu = round(mean(dist)), sdx = round(sd(dist)), median = round(median(dist)), rangex = round(range(dist)))
snpstats
```

Salmonella Infantis population structure - relative frequencies

```{r, include = FALSE}
# Enter MLST results

mlst <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/infantis/input_data/salmonellast_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'_'), "[", 1)

# Select columns of interest

s1 <- mlst %>%
          select(id2, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") %>%
                  rename(id = id2)

# st

s1$st <- ifelse(s1$ST == 32, "ST32", "Other STs")

data1 <- s1

# check the number of unique STs 

a <- data1 %>% select(ST) %>% drop_na() %>% distinct(ST) %>% count()
a

####################################################################

# Enter SISTR results

sistr <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/infantis/input_data/sistr_all.csv')

# QC the data 

skim(sistr)

# Generate the id column 

sistr$id <- sapply(strsplit(as.character(sistr$genome),'_'), "[", 1)

# Select columns 

sis1 <- sistr %>%
            select(id, cgmlst_ST, serovar_cgmlst) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") 

sis1$serovar <- ifelse(sis1$serovar_cgmlst == "Infantis", "Infantis",
                            "Other serovars")

data2 <- sis1

# check the number of unique cgmlts 

a <- data2 %>% select(cgmlst_ST) %>% drop_na() %>% distinct(cgmlst_ST) %>% count()
a

cgmlst_prop <- data2 %>%
              drop_na(cgmlst_ST) %>%
              mutate(cgmlst_ST = as.factor(cgmlst_ST)) %>%
              select(cgmlst_ST) %>%
              group_by(cgmlst_ST) %>%
              summarise(n = n()) %>%
              mutate(prop_cgmlst = n/sum(n)*100) %>%
              arrange(desc(prop_cgmlst))
############################################################

# Enter Baps 1-6 

baps <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/infantis/input_data/fastbaps_partition_baps_prior_l6.csv')

# Changing column names 

colnames(baps)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps %>% 
              select(id, baps_1)


data3 <- b1

# check the number of unique baps1 sub-groups

a <- data3 %>% select(baps_1) %>% drop_na() %>% distinct(baps_1) %>% count()
a
############################################################################

# Merge data1 and data2

data_1_2 <- left_join(data1, data2, on = "id")

# Merge data_1_2 with data_baps

data4 <- left_join(data_1_2, data3, on = "id")


data5 <- data4 %>% 
                select(id, st, cgmlst_ST, serovar, baps_1)

############################################################################

# Plot Serovar distribution

serovar_data <- data5 %>%
              drop_na(serovar) %>%
              select(serovar) %>%
              group_by(serovar) %>%
              summarise(n = n()) %>%
              mutate(prop_serovar = n/sum(n)*100) 


sero_plot <- ggplot(serovar_data, aes(x = prop_serovar, y = serovar, fill = serovar)) + xlim(0, 100) +
  theme_bw() + 
  theme(legend.position = "none") +
  ylab("") +
  xlab("Proportion") +
  theme(axis.text.y = element_text(size = 30)) +
  theme(axis.title.y = element_text(size = 30, face = "bold")) +
  theme(axis.title.x = element_text(size = 30, face = "bold")) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.7, size = 26)) +
  scale_fill_manual(values=c("darkblue", "gray")) +
   ggtitle("Serovar") +
  theme(plot.title = element_text(size = 40, face = "bold")) + 
  geom_col() +
  annotate(geom = "text", label = "n = 1,686 genomes", x = 50, y = 2, 
                                  color = "black", size = 8)
sero_plot

##################################################################################

# Plot ST distribution

st_data <- data5 %>%
              drop_na(serovar, st) %>%
              select(serovar, st) %>%
              group_by(serovar, st) %>%
              summarise(n = n()) %>%
              mutate(prop_st = n/sum(n)*100) %>%
              filter(serovar != "Other serovars")
st_data$st <- factor(st_data$st, levels=c("ST32", "Other STs"))


st_plot <- ggplot(st_data, aes(x = st, y = prop_st, fill = st)) + ylim(0, 105) +
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("") +
  ylab("Proportion") +
  theme(axis.text.y = element_text(size = 27)) +
  theme(axis.title.y = element_text(size = 30, face = "bold")) +
  theme(axis.title.x = element_text(size = 35, face = "bold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 30)) +
  scale_fill_manual(values=c("red", "beige")) +
  ggtitle("ST") +
  theme(plot.title = element_text(size = 40, face = "bold")) + 
  theme(strip.text.x = element_text(size = 20)) +
  geom_col() 
st_plot

#############################################################

# Plot cgMLST distribution

cgmlst_data <- data5 %>%
              drop_na(serovar, cgmlst_ST) %>%
              mutate(cgmlst_ST = as.factor(cgmlst_ST)) %>%
              select(serovar, cgmlst_ST) %>%
              group_by(serovar, cgmlst_ST) %>%
              summarise(n = n()) %>%
              mutate(prop_cgmlst = n/sum(n)*100) %>%
              arrange(serovar, desc(prop_cgmlst)) %>%
              filter(serovar != "Other serovars" & cgmlst_ST != "0" & prop_cgmlst > 2)

cgmlst_plot <- ggplot(cgmlst_data, aes(x = reorder(cgmlst_ST, -prop_cgmlst), y = prop_cgmlst, fill = cgmlst_ST)) + ylim(0, 100) +
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("cgMLST clones") +
  ylab("Proportion") +
  theme(axis.text.y = element_text(size = 25)) +
  theme(axis.title.y = element_text(size = 35, face = "bold")) +
  theme(axis.title.x = element_text(size = 35, face = "bold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 25)) +
  scale_fill_manual(values=c("red", "beige", "aliceblue", "darkgreen", "black", "brown", "orange", "gray")) +
  labs(fill="cgMLST clones") + 
  ggtitle("cgMLST") +
  theme(plot.title = element_text(size = 40, face = "bold")) + 
  geom_col() +
  annotate(geom = "text", label = "> 2%", x = 3, y = 90, 
                                  color = "black", size = 8)
cgmlst_plot

################################################################

# Plot baps1 through 6 distribution by serovar

baps_data <- data5 %>%
              drop_na(serovar, baps_1) %>%
              select(serovar, baps_1) %>%
              group_by(serovar, baps_1) %>%
              summarise(n = n()) %>%
              mutate(prop = n/sum(n)*100) %>%
              filter(serovar != "Other serovars")
baps_data$baps_1 <- factor(baps_data$baps_1, levels=c("3", "1"))

baps_plot <- ggplot(baps_data, aes(x = baps_1, y = prop, fill = baps_1)) + ylim(0, 105) +
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("") +
  ylab("Proportion") +
  theme(axis.text.y = element_text(size = 30)) +
  theme(axis.title.y = element_text(size = 35, face = "bold")) +
  theme(axis.title.x = element_text(size = 35, face = "bold")) +
  theme(axis.text.x = element_text(angle = 0, size = 30)) +
  scale_fill_manual(name = "Serovar", values=c("darkgreen", "brown")) +
  ggtitle("BAPS level 1") +
  theme(plot.title = element_text(size = 40, face = "bold")) + 
  theme(strip.text.x = element_text(size = 20)) +
  geom_col() 
baps_plot

fig4be  <- ggarrange(sero_plot, baps_plot, st_plot, cgmlst_plot,
                  nrow = 2, ncol = 2, widths = c(10, 10, 10, 10),
                heights = c(1, 1, 1, 1))
fig4be
```

Simpson diversity using ST, cgmlst, and baps 1-6 S. Infantis

```{r, include = FALSE}

mlst <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/infantis/input_data/salmonellast_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'_'), "[", 1)

# Select columns of interest

s1 <- mlst %>%
          select(id2, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") %>%
                  rename(id = id2)

data1 <- s1

####################################################################

# Enter SISTR results

sistr <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/infantis/input_data/sistr_all.csv')

# QC the data 

skim(sistr)

# Generate the id column 

sistr$id <- sapply(strsplit(as.character(sistr$genome),'_'), "[", 1)

# Select columns 

sis1 <- sistr %>%
            select(id, cgmlst_ST, serovar_cgmlst) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") %>%
              mutate(cgMLST = cgmlst_ST)

sis1$serovar <- ifelse(sis1$serovar_cgmlst == "Infantis", "Infantis",
                            "Other serovars")

data2 <- sis1

############################################################

# Enter Baps 1-6 

baps <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/infantis/input_data/fastbaps_partition_baps_prior_l6.csv')

# Changing column names 

colnames(baps)[1:7] <- c("id", "BAPS1", "BAPS2", "BAPS3", "BAPS4", "BAPS5", "BAPS6") 


data3 <- baps

# Merge 

data4 <- left_join(data1, data2, on = "id")

# Merge

data5 <- left_join(data3, data4, on = "id")

####################################################

# ST diversity 

st_div <- data5 %>%
              drop_na(serovar, ST) %>%
              mutate(ST = as.factor(ST)) %>%
              select(serovar, ST) %>%
              group_by(serovar, ST) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "ST")

######################################

# cgMLST diversity 

cgmlst_div <- data5 %>%
              drop_na(serovar, cgMLST) %>%
              mutate(cgMLST = as.factor(cgMLST)) %>%
              select(serovar, cgMLST) %>%
              group_by(serovar, cgMLST) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "cgMLST")

#####################################

# Baps based diversity 

# Baps1
baps1 <- data5 %>%
              select(serovar, BAPS1) %>%
              drop_na(serovar, BAPS1) %>%
              group_by(serovar, BAPS1) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS1")

# Baps2
baps2 <- data5 %>%
              select(serovar, BAPS2) %>%
              drop_na(serovar, BAPS2) %>%
              group_by(serovar, BAPS2) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS2")


# Baps3
baps3 <- data5 %>%
              select(serovar, BAPS3) %>%
              drop_na(serovar, BAPS3) %>%
              group_by(serovar, BAPS3) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS3")


# Baps4
baps4 <- data5 %>%
              select(serovar, BAPS4) %>%
              drop_na(serovar, BAPS4) %>%
              group_by(serovar, BAPS4) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS4")

# Baps5
baps5 <- data5 %>%
              select(serovar, BAPS5) %>%
              drop_na(serovar, BAPS5) %>%
              group_by(serovar, BAPS5) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS5")

# Baps6
baps6 <- data5 %>%
              select(serovar, BAPS6) %>%
              drop_na(serovar, BAPS6) %>%
              group_by(serovar, BAPS6) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS6")

# Concatenate all datasets

data6 <- rbind(st_div, cgmlst_div, baps1, baps2, baps3, baps4, baps5, baps6)
data6$strat <- factor(data6$strat, levels=c("ST", "cgMLST", "BAPS1", "BAPS2", "BAPS3",
                                              "BAPS4", "BAPS5", "BAPS6"))


# Plot Simpson's diversity analysis

fig4f <- ggplot(data6, aes(x = strat, y = value)) + ylim(0,1.05)+
  theme_bw() + 
  xlab("") +
  ylab("Index value") +
  theme(axis.text.y = element_text(size = 45)) +
  theme(axis.title.y = element_text(size = 50, face = "bold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 45)) +
  theme(axis.title.x = element_text(size = 26, face = "bold")) +
  theme(strip.text.x = element_text(size = 36))+
  ggtitle("") +
  theme(plot.title = element_text(size = 50, face = "bold")) + 
  theme(legend.text=element_text(size=20)) +
  geom_bar(stat = "identity", fill="darkblue") 
fig4f
```

Salmonella Newport phylogeny

```{r, include = FALSE}

# Enter the phylogeny for Typhimurium group1

tree <- read.tree("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/newport/input_data/newport_phylogeny.tree")

#############################################

# Enter Baps 1-6 

baps <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/newport/input_data/fastbaps_partition_baps_prior_l6.csv')

# Changing column names 

colnames(baps)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium//data/newport/input_data/salmonellast_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'_'), "[", 1)

# Select columns of interest

s1 <- mlst %>%
          select(id2, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") %>%
                  rename(id = id2)

# QC the data

skim(s1)

####################################################################

# Enter SISTR results

sistr <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium//data/newport/input_data/sistr_output.csv')

# QC the data 

skim(sistr)

# Generate the id column 

sistr$id <- sapply(strsplit(as.character(sistr$genome),'_'), "[", 1)

# Select columns 

sis1 <- sistr %>%
            select(id, cgmlst_ST, serovar_cgmlst) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") 

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mslt first

d1 <- left_join(b1, s1, on = "id")
d2 <- left_join(d1, sis1, on = "id")

# Create new columns for ST, cgmlst, and serovar

# cgmlst

a <- d2 %>%
  group_by(ST, cgmlst_ST) %>%
  count() %>%
  arrange(desc(n))


d2 <- mutate(d2, cgmlst = ifelse(cgmlst_ST %in% 1468400426, "cgMLST 1468400426",
                                     ifelse(cgmlst_ST %in% 88443731, "cgMLST 88443731", 
                                            ifelse(cgmlst_ST %in% 1271156802, "cgMLST 1271156802",
                                            "Other cgMLSTs"))))

# st

d2$st <- ifelse(d2$ST == 5, "ST5", 
                   ifelse(d2$ST == 45, "ST45", 
                          ifelse(d2$ST == 118, "ST118", "Other STs")))

# Replace NA in the ST column

d2 <- d2 %>% mutate(st = replace_na(st, "Other STs"))

# Create serovar column

#data1 <- data1 %>% mutate(serovar_cgmlst = replace_na(serovar_cgmlst, 0))
d2$serovar <- ifelse(d2$serovar_cgmlst == "Newport", "Newport",
                            "Other serovars")
                     
# Check to make sure the relevant columns are ok

table(d2$st)
table(d2$cgmlst)
table(d2$serovar)
table(d2$baps_1)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d3 <- d2 %>% 
                    select("id", "serovar", "baps_1", "st", "cgmlst") %>%
                    mutate(baps_1 = as.factor(baps_1))

# Make id the rowname or index

d3 <- column_to_rownames(d3, var = "id") 

# Create the tree

n1 <- ggtree(tree, layout = "circular") + xlim(-50, NA)


fig3a <- gheatmap(n1, d3, offset=.0, width=10, colnames = FALSE) +
  scale_fill_manual(values = c("gray", "darkblue",
                               "cornflowerblue", "coral", "darkgreen", "red",
                                 "brown", "darkseagreen3", "darkcyan", "orange",
                               "cadetblue4", "black", "darkblue", "purple", 
                               "darkgreen", "red", "black", "darkblue", "gray"),
                    breaks = c("Newport", "Other serovars",
                               "1", "2", "3", "4", "5", "6", "7", "8", "9",
                               "ST5", "ST45", "ST118", "Other STs",
                               "cgMLST 1468400426", "cgMLST 88443731", "cgMLST 1271156802", "Other cgMLSTs"),
                    name="Serovar (innermost) -> BAPS1 -> ST -> cgMLST (outermost)") +
    ggtitle(expression(bold(paste("H. Hierarchical population structure - ", bolditalic(S),". Newport")))) +
  theme(plot.title = element_text(size = 40, face = "bold"), 
        legend.title=element_text(size=24, face = "bold"), 
    legend.text=element_text(size=22))
fig3a
```

Salmonella Newport - distribution of SNPs 

```{r, include = FALSE}

# enter the snp data

data1 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/newport/input_data/distance_snp_sites_newport_columns.csv')

# filter out distance between itself

data2 <- data1 %>% filter_(~id_1 != id_2)

# distribution of snps 

summary(data2$dist)

# Generate quantiles

data3 <- data2 %>% 
             summarize(quartile = quantile(dist, probs = c(0.25, 0.5, 0.75, 1)))
data3

# get stats on snp dist

snpstats <- data2 %>% 
           summarize(mu = round(mean(dist)), sdx = round(sd(dist)), median = round(median(dist)), rangex = round(range(dist)))
snpstats

# Plot the distribution of SNPs

snpnew <- ggplot(data2, aes(x = dist)) +
  geom_histogram(bins = 10, fill = "darkblue") +
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("Number of SNPs") +
  ylab("Frequency") +
  theme(axis.text.y = element_text(size = 40)) +
  theme(axis.title.y = element_text(size = 45, face = "bold")) +
  theme(axis.title.x = element_text(size = 45, face = "bold")) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.7, size = 40)) +
  ggtitle("") +
  theme(plot.title = element_text(size = 50, face = "bold")) + 
  annotate(geom = "text", label = "Mean = 14,164", x = 7e+04, y = 6e+06, 
                                  color = "black", size = 18) +
  annotate(geom = "text", label = "Median = 24,676", x = 7e+04, y = 5e+06, 
                                  color = "black", size = 18) +
  annotate(geom = "text", label = "SD = 13,082", x = 7e+04, y = 4e+06, 
                                  color = "black", size = 18) +
  annotate(geom = "text", label = "Range (0-127,985)", x = 7e+04, y = 3e+06, 
                                  color = "black", size = 18) 
snpnew
```

Salmonella Newport population structure - relative frequencies

```{r, include = FALSE}

# Enter MLST results

mlst <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/newport/input_data/salmonellast_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'_'), "[", 1)

# Select columns of interest

s1 <- mlst %>%
          select(id2, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") %>%
                  rename(id = id2)

# st

s1$st <- ifelse(s1$ST == 5, "ST5", 
                   ifelse(s1$ST == 45, "ST45", 
                          ifelse(s1$ST == 118, "ST118", "Other STs")))

data1 <- s1

# check the number of unique STs

a <- data1 %>% select(ST) %>% drop_na() %>% distinct(ST) %>% count()
a

st_prop <- s1 %>%
              drop_na(st) %>%
              select(st) %>%
              group_by(st) %>%
              summarise(n = n()) %>%
              mutate(prop_st = n/sum(n)*100) %>%
              arrange(desc(prop_st))
####################################################################

# Enter SISTR results

sistr <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/newport/input_data/sistr_output.csv')

# QC the data 

skim(sistr)

# Generate the id column 

sistr$id <- sapply(strsplit(as.character(sistr$genome),'_'), "[", 1)

# Select columns 

sis1 <- sistr %>%
            select(id, cgmlst_ST, serovar_cgmlst) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") 

sis1$serovar <- ifelse(sis1$serovar_cgmlst == "Newport", "Newport",
                            "Other serovars")

data2 <- sis1

# check the number of unique baps_1 sub-groups

a <- data2 %>% select(cgmlst_ST) %>% drop_na() %>% distinct(cgmlst_ST) %>% count()
a
############################################################

# Enter Baps 1-6 

baps <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/newport/input_data/fastbaps_partition_baps_prior_l6.csv')

# Changing column names 

colnames(baps)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps %>% 
              select(id, baps_1)


data3 <- b1

# check the number of unique baps_1

a <- data3 %>% select(baps_1) %>% drop_na() %>% distinct(baps_1) %>% count()
a
############################################################################

# Merge data1 and data2

data_1_2 <- left_join(data1, data2, on = "id")

# Merge data_1_2 with data_baps

data4 <- left_join(data_1_2, data3, on = "id")


data5 <- data4 %>% 
                select(id, st, cgmlst_ST, serovar, baps_1)

############################################################################

# Plot Serovar distribution

serovar_data <- data5 %>%
              drop_na(serovar) %>%
              select(serovar) %>%
              group_by(serovar) %>%
              summarise(n = n()) %>%
              mutate(prop_serovar = n/sum(n)*100) 


sero_plot <- ggplot(serovar_data, aes(x = prop_serovar, y = serovar, fill = serovar)) + xlim(0, 100) +
  theme_bw() + 
  theme(legend.position = "none") +
  ylab("") +
  xlab("Proportion") +
  theme(axis.text.y = element_text(size = 30)) +
  theme(axis.title.y = element_text(size = 30, face = "bold")) +
  theme(axis.title.x = element_text(size = 30, face = "bold")) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.7, size = 26)) +
  scale_fill_manual(values=c("gray", "darkblue")) +
   ggtitle("Serovar") +
  theme(plot.title = element_text(size = 40, face = "bold")) + 
  geom_col() +
  annotate(geom = "text", label = "n = 2,365 genomes", x = 60, y = 2, 
                                  color = "black", size = 8)
sero_plot

##################################################################################

# Plot ST distribution

st_data <- data5 %>%
              drop_na(serovar, st) %>%
              select(serovar, st) %>%
              group_by(serovar, st) %>%
              summarise(n = n()) %>%
              mutate(prop_st = n/sum(n)*100) %>%
              filter(serovar != "Other serovars")
st_data$st <- factor(st_data$st, levels=c("ST118", "ST45", "ST5", "Other STs"))


st_plot <- ggplot(st_data, aes(x = st, y = prop_st, fill = st)) + ylim(0, 105) +
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("") +
  ylab("Proportion") +
    theme(axis.text.y = element_text(size = 28)) +
  theme(axis.title.y = element_text(size = 30, face = "bold")) +
  theme(axis.title.x = element_text(size = 30, face = "bold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 28)) +
  scale_fill_manual(values=c("purple", "darkblue", "black", "darkgreen")) +
  ggtitle("ST") +
  theme(plot.title = element_text(size = 40, face = "bold")) + 
  theme(strip.text.x = element_text(size = 20)) +
  geom_col() 
st_plot

#############################################################

# Plot cgMLST distribution

cgmlst_data <- data5 %>%
              drop_na(serovar, cgmlst_ST) %>%
              mutate(cgmlst_ST = as.factor(cgmlst_ST)) %>%
              select(serovar, cgmlst_ST) %>%
              group_by(serovar, cgmlst_ST) %>%
              summarise(n = n()) %>%
              mutate(prop_cgmlst = n/sum(n)*100) %>%
              arrange(serovar, desc(prop_cgmlst)) %>%
              filter(serovar != "Other serovars" & cgmlst_ST != "0" & prop_cgmlst > 2)

cgmlst_plot <- ggplot(cgmlst_data, aes(x = reorder(cgmlst_ST, -prop_cgmlst), y = prop_cgmlst, fill = cgmlst_ST)) + ylim(0, 100) +
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("cgMLST clones") +
  ylab("Proportion") +
      theme(axis.text.y = element_text(size = 28)) +
  theme(axis.title.y = element_text(size = 30, face = "bold")) +
  theme(axis.title.x = element_text(size = 30, face = "bold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 28)) +
  scale_fill_manual(values=c("black", "darkblue", "red", "darkgreen",
                          "orange", "brown")) +
  labs(fill="cgMLST clones") + 
  ggtitle("cgMLST") +
  theme(plot.title = element_text(size = 40, face = "bold")) + 
  geom_col() +
  annotate(geom = "text", label = "> 2%", x = 3, y = 90, 
                                  color = "black", size = 9)
cgmlst_plot

################################################################

# Plot baps1 through 6 distribution by serovar

baps_data <- data5 %>%
              drop_na(serovar, baps_1) %>%
              select(serovar, baps_1) %>%
              group_by(serovar, baps_1) %>%
              summarise(n = n()) %>%
              mutate(prop = n/sum(n)*100) %>%
              filter(serovar != "Other serovars")
baps_data$baps_1 <- factor(baps_data$baps_1, levels=c("8", "1", "4", "7", "6"))

baps_plot <- ggplot(baps_data, aes(x = baps_1, y = prop, fill = baps_1)) + ylim(0, 105) +
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("") +
  ylab("Proportion") +
    theme(axis.text.y = element_text(size = 30)) +
  theme(axis.title.y = element_text(size = 30, face = "bold")) +
  theme(axis.title.x = element_text(size = 30, face = "bold")) +
  theme(axis.text.x = element_text(angle = 0, size = 30)) +
  theme(legend.title=element_text(size=30, face = "bold"))+
  theme(legend.text=element_text(size=24)) +
  scale_fill_manual(name = "Serovar", values=c("orange", "cornflowerblue", "red", "darkcyan", "darkseagreen3")) +
  ggtitle("BAPS level 1") +
  theme(plot.title = element_text(size = 40, face = "bold")) + 
  theme(strip.text.x = element_text(size = 20)) +
  geom_col() 
baps_plot

fig3be  <- ggarrange(sero_plot, baps_plot, st_plot, cgmlst_plot,
                  nrow = 2, ncol = 2, widths = c(10, 10, 10, 10),
                heights = c(1, 1, 1, 1))
fig3be
```


Plot Simpson diversity using ST, cgmlst, and baps 1-6 S. Newport

```{r, include = FALSE}

mlst <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/newport/input_data/salmonellast_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'_'), "[", 1)

# Select columns of interest

s1 <- mlst %>%
          select(id2, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") %>%
                  rename(id = id2)

data1 <- s1

####################################################################

# Enter SISTR results

sistr <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/newport/input_data/sistr_output.csv')

# QC the data 

skim(sistr)

# Generate the id column 

sistr$id <- sapply(strsplit(as.character(sistr$genome),'_'), "[", 1)

# Select columns 

sis1 <- sistr %>%
            select(id, cgmlst_ST, serovar_cgmlst) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") %>%
              mutate(cgMLST = cgmlst_ST)

sis1$serovar <- ifelse(sis1$serovar_cgmlst == "Newport", "Newport",
                            "Other serovars")

data2 <- sis1

############################################################

# Enter Baps 1-6 

baps <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/newport/input_data/fastbaps_partition_baps_prior_l6.csv')

# Changing column names 

colnames(baps)[1:7] <- c("id", "BAPS1", "BAPS2", "BAPS3", "BAPS4", "BAPS5", "BAPS6") 


data3 <- baps

# Merge 

data4 <- left_join(data1, data2, on = "id")

# Merge

data5 <- left_join(data3, data4, on = "id")

####################################################

# ST diversity 

st_div <- data5 %>%
              drop_na(serovar, ST) %>%
              mutate(ST = as.factor(ST)) %>%
              select(serovar, ST) %>%
              group_by(serovar, ST) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "ST")

######################################

# cgMLST diversity 

cgmlst_div <- data5 %>%
              drop_na(serovar, cgMLST) %>%
              mutate(cgMLST = as.factor(cgMLST)) %>%
              select(serovar, cgMLST) %>%
              group_by(serovar, cgMLST) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "cgMLST")

#####################################

# Baps based diversity 

# Baps1
baps1 <- data5 %>%
              select(serovar, BAPS1) %>%
              drop_na(serovar, BAPS1) %>%
              group_by(serovar, BAPS1) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS1")

# Baps2
baps2 <- data5 %>%
              select(serovar, BAPS2) %>%
              drop_na(serovar, BAPS2) %>%
              group_by(serovar, BAPS2) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS2")


# Baps3
baps3 <- data5 %>%
              select(serovar, BAPS3) %>%
              drop_na(serovar, BAPS3) %>%
              group_by(serovar, BAPS3) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS3")


# Baps4
baps4 <- data5 %>%
              select(serovar, BAPS4) %>%
              drop_na(serovar, BAPS4) %>%
              group_by(serovar, BAPS4) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS4")

# Baps5
baps5 <- data5 %>%
              select(serovar, BAPS5) %>%
              drop_na(serovar, BAPS5) %>%
              group_by(serovar, BAPS5) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS5")

# Baps6
baps6 <- data5 %>%
              select(serovar, BAPS6) %>%
              drop_na(serovar, BAPS6) %>%
              group_by(serovar, BAPS6) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS6")

# Concatenate all datasets

data6 <- rbind(st_div, cgmlst_div, baps1, baps2, baps3, baps4, baps5, baps6)
data6$strat <- factor(data6$strat, levels=c("ST", "cgMLST", "BAPS1", "BAPS2", "BAPS3",
                                              "BAPS4", "BAPS5", "BAPS6"))


# Plot Simpson's diversity analysis

fig3f <- ggplot(data6, aes(x = strat, y = value)) + ylim(0,1.05)+
  theme_bw() + 
  xlab("") +
  ylab("Index value") +
  theme(axis.text.y = element_text(size = 45)) +
  theme(axis.title.y = element_text(size = 50, face = "bold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 45)) +
  theme(axis.title.x = element_text(size = 26, face = "bold")) +
  theme(strip.text.x = element_text(size = 36))+
  ggtitle("") +
  theme(plot.title = element_text(size = 50, face = "bold")) + 
  theme(legend.text=element_text(size=20)) +
  geom_bar(stat = "identity", fill="darkblue") 
fig3f
```


Salmonella Typhimurium phylogeny with metadata using group1 as a representative 

```{r, include = FALSE}

# Enter the phylogeny for Typhimurium group1

tree_group_1 <- read.tree("~/Documents/frontiers_paper_salmonella_newport_typhimurium//data/typhimurium/group_1/input_data/typhimurium_phylogeny_gr1.tree")

#############################################

# Enter Baps 1-6 for Typhimurium group1

baps_group_1 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium//data/typhimurium/group_1/input_data/roary_output_gr1.fastbaps_l6.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium//data/typhimurium/global_input_data/salmonellast_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'/'), "[", 3)
s1 <- mlst %>% separate(id2, c("id", "day"), sep = "\\.")

# Select columns of interest

s2 <- s1 %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s2)

####################################################################

# Enter SISTR results

sistr <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium//data/typhimurium/global_input_data/sistr_all.csv')

# QC the data 

skim(sistr)

# Select columns 

sis1 <- sistr %>%
            select(genome, cgmlst_ST, serovar_cgmlst) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") %>%
                rename(id = genome)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mslt first

d1 <- left_join(b1, s2, on = "id") # merge mlst and baps
d2 <- left_join(d1, sis1, on = "id") # merge d1 with sistr 

# Create new columns for ST, cgmlst, and serovar

# cgmlst

d2 <- mutate(d2, cgmlst = ifelse(cgmlst_ST %in% 1652656062, "cgMLST 1652656062",
                                     ifelse(cgmlst_ST %in% 860079270, "cgMLST 860079270", "Other cgMLSTs")))

# st

d2$st <- ifelse(d2$ST == 19, "ST19", 
                   ifelse(d2$ST == 34, "ST34", 
                          ifelse(d2$ST == 36, "ST36",
                                 ifelse(d2$ST == 313, "ST313", "Other STs"))))

# Replace NA in the ST column

d2 <- d2 %>% mutate(st = replace_na(st, "Other STs"))

# Create serovar column

#data1 <- data1 %>% mutate(serovar_cgmlst = replace_na(serovar_cgmlst, 0))
d2$serovar <- ifelse(d2$serovar_cgmlst == "Typhimurium", "Biphasic",
                     ifelse(d2$serovar_cgmlst == "I 4,[5],12:i:-", "Monophasic",
                            "Other serovars"))
                     
# Check to make sure the relevant columns are ok

table(d2$st)
table(d2$cgmlst)
table(d2$serovar)
table(d2$baps_1)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d3 <- d2 %>% 
                    select("id", "serovar", "baps_1", "st", "cgmlst") %>%
                    mutate(baps_1 = as.factor(baps_1))

# Make id the rowname or index

d3 <- column_to_rownames(d3, var = "id") 

# Create the tree

t1 <- ggtree(tree_group_1, layout = "circular") + xlim(-50, NA)


fig2a <- gheatmap(t1, d3, offset=.0, width=10, colnames = FALSE) +
  scale_fill_manual(values = c("gray", "darkblue",
                               "cornflowerblue", "coral", "darkgreen", "red",
                                 "red", "darkseagreen3", "darkcyan", "darksalmon",
                               "cadetblue4", "black", "darkblue", "purple", 
                               "red", "darkblue", "gray"),
                    breaks = c("Biphasic", "Monophasic", "Other serovars",
                               "1", "2", "3", "4", "5", "6",
                               "ST19", "ST34", "ST36", "ST313", "Other STs",
                               "cgMLST 1652656062", "cgMLST 860079270", "Other cgMLSTs"),
                    name= "Serovar (innermost) -> BAPS1 -> ST -> cgMLST (outermost)") +
  ggtitle(expression(bold(paste("O. Hierarchical population structure - ", bolditalic(S),". Typhimurium")))) +
  theme(plot.title = element_text(size = 40, face = "bold"), 
        legend.title=element_text(size=24, face = "bold"), 
    legend.text=element_text(size=22))
fig2a
```

Salmonella Typhimurium - distribution of SNPs 

```{r, include = FALSE}

# enter the snp data

data1 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_1/input_data/distance_snp_sites_typhimurium_1_columns.csv')

data2 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_2/input_data/distance_snp_sites_typhimurium_2_columns.csv')

data3 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_3/input_data/distance_snp_sites_typhimurium_3_columns.csv')

data4 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_4/input_data/distance_snp_sites_typhimurium_4_columns.csv')

data5 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_5/input_data/distance_snp_sites_typhimurium_5_columns.csv')

data6 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_6/input_data/distance_snp_sites_typhimurium_6_columns.csv')

data7 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_7/input_data/distance_snp_sites_typhimurium_7_columns.csv')

data8 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_8/input_data/distance_snp_sites_typhimurium_8_columns.csv')

data9 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_9/input_data/distance_snp_sites_typhimurium_9_columns.csv')

data10 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_10/input_data/distance_snp_sites_typhimurium_10_columns.csv')

data11 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_11/input_data/distance_snp_sites_typhimurium_11_columns.csv')

data12 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_12/input_data/distance_snp_sites_typhimurium_12_columns.csv')

data13 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_13/input_data/distance_snp_sites_typhimurium_13_columns.csv')

data14 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_14/input_data/distance_snp_sites_typhimurium_14_columns.csv')

data15 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_15/input_data/distance_snp_sites_typhimurium_15_columns.csv')

data16 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_16/input_data/distance_snp_sites_typhimurium_16_columns.csv')

data17 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_17/input_data/distance_snp_sites_typhimurium_17_columns.csv')

data18 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_18/input_data/distance_snp_sites_typhimurium_18_columns.csv')

data19 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_19/input_data/distance_snp_sites_typhimurium_19_columns.csv')

data20 <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_20/input_data/distance_snp_sites_typhimurium_20_columns.csv')

# concatenate the datasets

data21 <- bind_rows(data1, data2, data3, data4, data5, data6, data7, data8, data9,
                    data10, data11, data12, data13, data14, data15, data16, data17, data18,
                    data19, data20)

# filter out distance between itself

data22 <- data21 %>% filter_(~id_1 != id_2)

# distribution of snps 

summary(data22$dist)

# Generate quantiles

data23 <- data22 %>% 
             summarize(quartile = quantile(dist, probs = c(0.25, 0.5, 0.75, 1)))
data23

# get stats on snp dist

snpstats <- data22 %>% 
           summarize(mu = round(mean(dist)), sdx = round(sd(dist)), median = round(median(dist)), rangex = round(range(dist)))
snpstats

# Plot the distribution of SNPs

snptyphi <- ggplot(data22, aes(x = dist)) +
  geom_histogram(bins = 10, fill = "darkblue") +
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("Number of SNPs") +
  ylab("Frequency") +
  theme(axis.text.y = element_text(size = 40)) +
  theme(axis.title.y = element_text(size = 45, face = "bold")) +
  theme(axis.title.x = element_text(size = 45, face = "bold")) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.7, size = 40)) +
  ggtitle("") +
  theme(plot.title = element_text(size = 50, face = "bold")) + 
  annotate(geom = "text", label = "Mean = 1,474", x = 50000, y = 2e+07, 
                                  color = "black", size = 18) +
  annotate(geom = "text", label = "Median = 281", x = 50000, y = 1.5e+07, 
                                  color = "black", size = 18) +
  annotate(geom = "text", label = "SD = 4,547", x = 50000, y = 1e+07, 
                                  color = "black", size = 18) +
  annotate(geom = "text", label = "Range (0-97,140)", x = 50000, y = 5e+06, 
                                  color = "black", size = 18) 
snptyphi
```

Salmonella Typhimurium population structure - relative frequencies

```{r, include = FALSE}

# Enter MLST results

mlst <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium//data/typhimurium/global_input_data/salmonellast_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'/'), "[", 3)
s1 <- mlst %>% separate(id2, c("id", "day"), sep = "\\.")

# Select columns of interest

s2 <- s1 %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s2)

# st

s2$st <- ifelse(s2$ST == 19, "ST19", 
                   ifelse(s2$ST == 34, "ST34", 
                          ifelse(s2$ST == 36, "ST36",
                                 ifelse(s2$ST == 313, "ST313", "Other STs"))))

# Create data1 

data1 <- s2 

# check the number of unique STs

a <- data1 %>% select(ST) %>% drop_na() %>% distinct(ST) %>% count()
a

####################################################################

# Enter SISTR results

sistr <- read_csv('~/Documents/frontiers_paper_salmonella_newport_typhimurium//data/typhimurium/global_input_data/sistr_all.csv')

# QC the data 

skim(sistr)

# Select columns 

sis1 <- sistr %>%
            select(genome, cgmlst_ST, serovar_cgmlst) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?") %>%
                rename(id = genome)

# Create serovar column

#data1 <- data1 %>% mutate(serovar_cgmlst = replace_na(serovar_cgmlst, 0))
sis1$serovar <- ifelse(sis1$serovar_cgmlst == "Typhimurium", "Biphasic",
                     ifelse(sis1$serovar_cgmlst == "I 4,[5],12:i:-", "Monophasic",
                            "Other serovars"))

# Create data2 and drop rows with missing values

data2 <- sis1 

# check the number of unique cgmlsts

a <- data2 %>% select(cgmlst_ST) %>% drop_na() %>% distinct(cgmlst_ST) %>% count()
a

##########################################################################

# Enter group1-Baps

# Enter the data, create a new column for group, and generate the id column
data3 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/group_1/input_data/roary_output_gr1.fastbaps_l6.csv")
data3$group <- "group_1"
colnames(data3)[1] <- "id"

###########################################################################

# Enter group2-Baps

# Enter the data, create a new column for group, and generate the id column

data4 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr2.fastbaps_l6.csv")
data4$group <- "group_2"
colnames(data4)[1] <- "id"

###########################################################################

# Enter group3-Baps
# Enter the data, create a new column for group, and generate the id column

data5 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr3.fastbaps_l6.csv")
data5$group <- "group_3"
colnames(data5)[1] <- "id"

###########################################################################

# Enter group4-Baps
# Enter the data, create a new column for group, and generate the id column

data6 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr4.fastbaps_l6.csv")
data6$group <- "group_4"
colnames(data6)[1] <- "id"

###########################################################################

# Enter group5-Baps
# Enter the data, create a new column for group, and generate the id column

data7 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr5.fastbaps_l6.csv")
data7$group <- "group_5"
colnames(data7)[1] <- "id"

###########################################################################

# Enter group6-Baps
# Enter the data, create a new column for group, and generate the id column

data8 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr6.fastbaps_l6.csv")
data8$group <- "group_6"
colnames(data8)[1] <- "id"

###########################################################################

# Enter group7-Baps
# Enter the data, create a new column for group, and generate the id column

data9 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr7.fastbaps_l6.csv")
data9$group <- "group_7"
colnames(data9)[1] <- "id"

###########################################################################

# Enter group8-Baps
# Enter the data, create a new column for group, and generate the id column

data10 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr8.fastbaps_l6.csv")
data10$group <- "group_8"
colnames(data10)[1] <- "id"

###########################################################################

# Enter group9-Baps
# Enter the data, create a new column for group, and generate the id column

data11 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr9.fastbaps_l6.csv")
data11$group <- "group_9"
colnames(data11)[1] <- "id"

###########################################################################

# Enter group10-Baps
# Enter the data, create a new column for group, and generate the id column

data12 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr10.fastbaps_l6.csv")
data12$group <- "group_10"
colnames(data12)[1] <- "id"

###########################################################################

# Enter group11-Baps
# Enter the data, create a new column for group, and generate the id column

data13 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr11.fastbaps_l6.csv")
data13$group <- "group_11"
colnames(data13)[1] <- "id"

###########################################################################

# Enter group12-Baps
# Enter the data, create a new column for group, and generate the id column

data14 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr12.fastbaps_l6.csv")
data14$group <- "group_12"
colnames(data14)[1] <- "id"

###########################################################################

# Enter group13-Baps
# Enter the data, create a new column for group, and generate the id column

data15 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr13.fastbaps_l6.csv")
data15$group <- "group_13"
colnames(data15)[1] <- "id"

###########################################################################

# Enter group14-Baps
# Enter the data, create a new column for group, and generate the id column

data16 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr14.fastbaps_l6.csv")
data16$group <- "group_14"
colnames(data16)[1] <- "id"

###########################################################################

# Enter group15-Baps
# Enter the data, create a new column for group, and generate the id column

data17 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr15.fastbaps_l6.csv")
data17$group <- "group_15"
colnames(data17)[1] <- "id"

###########################################################################

# Enter group16-Baps
# Enter the data, create a new column for group, and generate the id column

data18 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr16.fastbaps_l6.csv")
data18$group <- "group_16"
colnames(data18)[1] <- "id"

###########################################################################

# Enter group17-Baps
# Enter the data, create a new column for group, and generate the id column

data19 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr17.fastbaps_l6.csv")
data19$group <- "group_17"
colnames(data19)[1] <- "id"

###########################################################################

# Enter group18-Baps
# Enter the data, create a new column for group, and generate the id column

data20 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr18.fastbaps_l6.csv")
data20$group <- "group_18"
colnames(data20)[1] <- "id"

###########################################################################

# Enter group19-Baps
# Enter the data, create a new column for group, and generate the id column

data21 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr19.fastbaps_l6.csv")
data21$group <- "group_19"
colnames(data21)[1] <- "id"

###########################################################################

# Enter group20-Baps
# Enter the data, create a new column for group, and generate the id column

data22 <- read_csv("~/Documents/frontiers_paper_salmonella_newport_typhimurium/data/typhimurium/global_input_data/roary_output_gr20.fastbaps_l6.csv")
data22$group <- "group_20"
colnames(data22)[1] <- "id"

###########################################################################

# First merge all Baps datasets 

data_baps <- bind_rows(data3, data4, data5, data6, data7, data8,
                       data9, data10, data11, data12, data13, data14, 
                       data15, data16, data17, data18, data19, data20, 
                       data21, data22)

colnames(data_baps)[2:7] <- c("baps_1", "baps_2", "baps_3", "baps_4", "baps_5", "baps_6")

# check the number of unique baps_1  sub-groups

a <- data_baps %>% select(baps_1) %>% drop_na() %>% distinct(baps_1) %>% count()
a
############################################################################

# Merge data1 and data2

data_1_2 <- left_join(data1, data2, on = "id")

# Merge data_1_2 with data_baps

data23 <- left_join(data_1_2, data_baps, on = "id")
glimpse(data23)

data23b <- data23 %>% 
                select(id, group, st, cgmlst_ST, serovar, baps_1)
############################################################################

# baps1 between monophasic vs biphasic

a <- data23 %>% select(serovar, baps_1) %>% drop_na() %>% group_by(serovar, baps_1) %>% distinct()
b <- a %>% filter(serovar == "Biphasic") %>% count() 
m <- a %>% filter(serovar == "Monophasic") %>% count() 
o <- a %>% filter(serovar == "Other serovars") %>% count() 

# st between monophasic vs biphasic

a <- data23 %>% select(serovar, ST) %>% drop_na() %>% group_by(serovar, ST) %>% distinct()
b <- a %>% filter(serovar == "Biphasic") %>% count() 
m <- a %>% filter(serovar == "Monophasic") %>% count() 
o <- a %>% filter(serovar == "Other serovars") %>% count() 

# cgmlst between monophasic vs biphasic

a <- data23 %>% select(serovar, cgmlst_ST) %>% drop_na() %>% group_by(serovar, cgmlst_ST) %>% distinct()
b <- a %>% filter(serovar == "Biphasic") %>% count() 
m <- a %>% filter(serovar == "Monophasic") %>% count() 
o <- a %>% filter(serovar == "Other serovars") %>% count() 

############################################################################

# Plot Serovar distribution

serovar_data <- data23b %>%
              drop_na(serovar) %>%
              select(group, serovar) %>%
              group_by(group, serovar) %>%
              summarise(n = n()) %>%
              mutate(prop_serovar = n/sum(n)*100) 
serovar_data$serovar <- factor(serovar_data$serovar, levels=c("Biphasic",
                                                              "Monophasic",
                                                              "Other serovars"))

sero_plot <- ggplot(serovar_data, aes(x = prop_serovar, y = serovar, fill = serovar)) + xlim(0, 100) +
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("Proportion") +
  ylab("") +
theme(axis.text.y = element_text(size = 35)) +
  theme(axis.title.y = element_text(size = 30, face = "bold")) +
  theme(axis.title.x = element_text(size = 30, face = "bold")) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.7, size = 26)) +
  scale_fill_manual(values=c("gray", "darkblue", "lightblue")) +
  ggtitle("Serovar") +
  theme(plot.title = element_text(size = 40, face = "bold")) + 
   #ggtitle(expression(bold(paste("B. ", bolditalic(S),". Typhimurium", " - Serovar")))) +
  #ggtitle(label = "B.",
  #subtitle = "Serovar") +
  #theme(plot.title = element_text(size = 50, face = "bold"),
        #plot.subtitle = element_text(size = 26, face = "bold")) + 
  geom_boxplot() +
  annotate(geom = "text", label = "n = 21,534 genomes", x = 60, y = 3, 
                                  color = "black", size = 7)
sero_plot

##################################################################################

# Plot ST distribution

st_data <- data23b %>%
              drop_na(serovar, st) %>%
              select(group, serovar, st) %>%
              group_by(group, serovar, st) %>%
              summarise(n = n()) %>%
              mutate(prop_st = n/sum(n)*100) %>%
              filter(serovar != "Other serovars")
st_data$st <- factor(st_data$st, levels=c("ST19", "ST34", "ST36", "ST313", "Other STs"))

# Calculate a chi-square p-value for independence 
st_data2 <- data23b %>%
              drop_na(serovar, st)
chisq_st <- chisq.test(table(st_data2$serovar,st_data2$st))
chisq_st

st_plot <- ggplot(st_data, aes(x = st, y = prop_st, fill = st)) + ylim(0, 105) +
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("") +
  ylab("Proportion") +
theme(axis.text.y = element_text(size = 22)) +
  theme(axis.title.y = element_text(size = 30, face = "bold")) +
  theme(axis.title.x = element_text(size = 35, face = "bold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 25)) +
  theme(legend.title=element_text(size=30, face = "bold"))+
  theme(legend.text=element_text(size=24)) +
  scale_fill_manual(values=c("darkgreen", "darkblue", "gray", "red", "black")) +
  ggtitle("ST") +
  theme(plot.title = element_text(size = 40, face = "bold")) + 
  theme(strip.text.x = element_text(size = 20)) +
  geom_boxplot() +
  annotate(geom = "text", label = "italic(p) < 2.2e-16", x = 3.5, y = 75, 
                                  color = "black", size = 7, parse = TRUE) +
  facet_wrap(~ serovar)
st_plot

#############################################################

# Plot cgMLST distribution

cgmlst_data <- data23b %>%
              drop_na(serovar, cgmlst_ST) %>%
              mutate(cgmlst_ST = as.factor(cgmlst_ST)) %>%
              select(group, serovar, cgmlst_ST) %>%
              group_by(group, serovar, cgmlst_ST) %>%
              summarise(n = n()) %>%
              mutate(prop_cgmlst = n/sum(n)*100) %>%
              arrange(group, serovar, desc(prop_cgmlst)) %>%
              filter(serovar != "Other serovars" & cgmlst_ST != "0" & prop_cgmlst > 2)

cgmlst_plot <- ggplot(cgmlst_data, aes(x = reorder(cgmlst_ST, -prop_cgmlst), y = prop_cgmlst, fill = cgmlst_ST)) + ylim(0, 100) +
  theme_bw() + 
  theme(legend.position = "none") +
  xlab("cgMLST clones") +
  ylab("Proportion") +
theme(axis.text.y = element_text(size = 22)) +
  theme(axis.title.y = element_text(size = 30, face = "bold")) +
  theme(axis.title.x = element_text(size = 35, face = "bold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 20)) +
  theme(strip.text.x = element_text(size = 25)) +
  scale_fill_manual(values=c("red", "darkblue", "gray", "darkgreen",
                          "orange", "yellow", "purple", "white", "black", "brown")) +
  #scale_fill_gradient2(name = "cgmlst_ST", 
                        #low = "blue", high = "red") +
  labs(fill="cgMLST clones") + 
  ggtitle("cgMLST") +
  theme(plot.title = element_text(size = 40, face = "bold")) + 
  geom_boxplot(position = position_dodge(0.9)) +
  annotate(geom = "text", label = "> 2%", x = 3, y = 75, 
                                  color = "black", size = 8) +
  facet_wrap(~serovar)
cgmlst_plot

################################################################

# Plot baps1 through 6 distribution by serovar

baps_data <- data23b %>%
              drop_na(serovar, baps_1) %>%
              select(group, serovar, baps_1) %>%
              group_by(group, serovar, baps_1) %>%
              summarise(n = n()) %>%
              mutate(prop = n/sum(n)*100) %>%
              filter(serovar != "Other serovars")
baps_data$baps_1 <- factor(baps_data$baps_1, levels=c("1", "2", "3", "4", "5", "6"))

baps_plot <- ggplot(baps_data, aes(x = baps_1, y = prop, fill = serovar)) + ylim(0, 105) +
  theme_bw() + 
  #theme(legend.position = "none") +
  xlab("") +
  ylab("Proportion") +
theme(axis.text.y = element_text(size = 30)) +
  theme(axis.title.y = element_text(size = 30, face = "bold")) +
  theme(axis.title.x = element_text(size = 35, face = "bold")) +
  theme(axis.text.x = element_text(angle = 0, size = 28)) +
  scale_fill_manual(name = "Serovar", values=c("gray", "darkblue")) +
  ggtitle("BAPS level 1") +
  theme(plot.title = element_text(size = 40, face = "bold")) + 
  theme(strip.text.x = element_text(size = 20)) +
  theme(legend.title=element_text(size=30, face = "bold")) +
  theme(legend.text=element_text(size=24)) +
  geom_boxplot() 
baps_plot

fig2be  <- ggarrange(sero_plot, baps_plot, st_plot, cgmlst_plot,
                  nrow = 2, ncol = 2, widths = c(10, 10, 10, 10),
                heights = c(1, 1, 1, 1))
fig2be
```


Plot Genomic diversity analysis at the ST, cgMLST, and BAPS1-6 levels for S. Typhimurium

```{r, include = FALSE}

# ST diversity 

st_div <- data23 %>%
              drop_na(ST, serovar) %>%
              mutate(ST = as.factor(ST)) %>%
              select(group, serovar, ST) %>%
              group_by(group, serovar, ST) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(group, serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("group", "serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "ST")

# Compute t-test
simpttest <- t.test(value ~ serovar, data = st_div)
simpttest

######################################

# cgMLST diversity 

cgmlst_div <- data23 %>%
              drop_na(cgmlst_ST, serovar) %>%
              mutate(cgmlst_ST = as.factor(cgmlst_ST)) %>%
              select(group, serovar, cgmlst_ST) %>%
              group_by(group, serovar, cgmlst_ST) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(group, serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("group", "serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "cgMLST")

# Compute t-test
simpttest <- t.test(value ~ serovar, data = cgmlst_div)
simpttest

#####################################

# Baps based diversity 

# Baps1
baps1 <- data23 %>%
              select(group, serovar, baps_1) %>%
              drop_na(group, serovar, baps_1) %>%
              group_by(group, serovar, baps_1) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(group, serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("group", "serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS1")

# Compute t-test

b1simp <- t.test(value ~ serovar, data = baps1)
b1simp

# Baps2
baps2 <- data23 %>%
              select(group, serovar, baps_2) %>%
              drop_na(group, serovar, baps_2) %>%
              group_by(group, serovar, baps_2) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(group, serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("group", "serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS2")


# Compute t-test
b2simp <- t.test(value ~ serovar, data = baps2)
b2simp

# Baps3
baps3 <- data23 %>%
              select(group, serovar, baps_3) %>%
              drop_na(group, serovar, baps_3) %>%
              group_by(group, serovar, baps_3) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(group, serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("group", "serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS3")

# Compute t-test

b3simp <- t.test(value ~ serovar, data = baps3)
b3simp

# Baps4
baps4 <- data23 %>%
              select(group, serovar, baps_4) %>%
              drop_na(group, serovar, baps_4) %>%
              group_by(group, serovar, baps_4) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(group, serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("group", "serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS4")

# Compute t-test

b4simp <- t.test(value ~ serovar, data = baps4)
b4simp

# Baps5
baps5 <- data23 %>%
              select(group, serovar, baps_5) %>%
              drop_na(group, serovar, baps_5) %>%
              group_by(group, serovar, baps_5) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(group, serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("group", "serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS5")

# Compute t-test

b5simp <- t.test(value ~ serovar, data = baps5)
b5simp

# Baps6
baps6 <- data23 %>%
              select(group, serovar, baps_6) %>%
              drop_na(group, serovar, baps_6) %>%
              group_by(group, serovar, baps_6) %>%
              summarise(n = n()) %>%
              mutate(simpson = diversity(n, "simpson")) %>%
              filter(serovar != "Other serovars") %>%
              group_by(group, serovar) %>%
              summarise(simpson = mean(simpson)) %>%
              melt(id.vars=c("group", "serovar"), measure.vars="simpson",
                    variable.name="index", value.name="value") %>%
              mutate(strat = "BAPS6")

# Compute t-test

b6simp <- t.test(value ~ serovar, data = baps6)
b6simp


# Concatenate all datasets

data24 <- rbind(st_div, cgmlst_div, baps1, baps2, baps3, baps4, baps5, baps6)
data24$strat <- factor(data24$strat, levels=c("ST", "cgMLST", "BAPS1", "BAPS2", "BAPS3",
                                              "BAPS4", "BAPS5", "BAPS6"))


# Plot Simpson's diversity analysis

fig2f <- ggplot(data24, aes(x = strat, y = value, fill = serovar)) + ylim(0,1.05)+
  theme_bw() + 
  xlab("") +
  ylab("Index value") +
  theme(axis.text.y = element_text(size = 45)) +
  theme(axis.title.y = element_text(size = 50, face = "bold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 45)) +
  theme(axis.title.x = element_text(size = 26, face = "bold")) +
  theme(strip.text.x = element_text(size = 36))+
  scale_fill_manual(values=c("gray", "darkblue")) + 
  ggtitle("") +
  theme(plot.title = element_text(size = 50, face = "bold")) + 
    theme(legend.title=element_text(size=40, face = "bold")) +
  theme(legend.text=element_text(size=30)) +
  labs(fill='Serovar') +
  geom_boxplot() +
 annotate(geom = "text", label = "****", x = 1, y = 0.5, 
                                  color = "black", size = 6) +
    annotate(geom = "text", label = "****", x = 2, y = 1.02, 
                                  color = "black", size = 6) +
    annotate(geom = "text", label = "****", x = 3, y = 0.25, 
                                  color = "black", size = 6) +
    annotate(geom = "text", label = "****", x = 4, y = 1, 
                                  color = "black", size = 6) +
    annotate(geom = "text", label = "****", x = 5, y = 1, 
                                  color = "black", size = 6) +
    annotate(geom = "text", label = "****", x = 6, y = 1, 
                                  color = "black", size = 6) +
   annotate(geom = "text", label = "****", x = 7, y = 1.02, 
                                  color = "black", size = 6) +
    annotate(geom = "text", label = "***", x = 8, y = 1.02, 
                                  color = "black", size = 6) +
  geom_vline(xintercept = 1.5) +
  geom_vline(xintercept = 2.5) +
  geom_vline(xintercept = 3.5) +
  geom_vline(xintercept = 4.5) +
  geom_vline(xintercept = 5.5) +
  geom_vline(xintercept = 6.5) +
  geom_vline(xintercept = 7.5) 
fig2f
```

